---
title: "Visualizing and validating twig models"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Visualizing and validating twig models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4.5)
library(twig)
set.seed(1)
```

`twig` ships a small set of helpers for **validating** a model before you run it
and for **visualizing** its structure and results. This vignette walks through
them on a compact 3-state Markov model and a short decision tree.

## A Markov model

```{r model}
mytwig <- twig() +
  decisions(names = c(StandardOfCare, Intervention)) +
  states(names = c(Healthy, Sick, Dead), init_probs = c(1, 0, 0)) +
  event(name = transition,
        options = c(progress, die, none),
        probs   = c(pProgress, pDie, leftover),
        transitions = c(Sick, Dead, stay)) +
  payoffs(names = c(cost, qaly), discount_rates = c(0.03, 0.03))

# Vectorized model functions
pProgress <- function(state, decision, rr) {
  0.10 * (state == "Healthy") * ifelse(decision == "Intervention", rr, 1)
}
pDie  <- function(state) 0.01 * (state == "Healthy") + 0.05 * (state == "Sick")
cost  <- function(state, decision) {
  (state == "Healthy") * 1000 + (state == "Sick") * 4000 +
    (decision == "Intervention") * 2000
}
qaly  <- function(state) 1.0 * (state == "Healthy") + 0.7 * (state == "Sick")
```

## Validate before you run

`validate_twig()` is a pre-flight check: it confirms that every probability and
payoff function referenced by the model is defined and that its arguments are
core arguments (`decision`, `state`, `cycle`, `cycle_in_state`) or names in your
`params`.

```{r validate}
params_point <- list(rr = 0.5)
validate_twig(mytwig, params = params_point)
```

## Cohort trace

Run the model with `verbose = TRUE` to obtain the cohort trace, then plot the
share of the cohort in each state over time with `plot_trace()`.

```{r trace}
res_v <- run_twig(mytwig, params = params_point, n_cycles = 30,
                  verbose = TRUE, progress_bar = FALSE)
plot_trace(res_v)
```

## State-transition diagram

`draw_transitions()` turns the realized transition-probability matrix into a
directed graph. It generates Graphviz DOT in pure R; rendering uses the
suggested **DiagrammeR** package.

```{r transitions-dot}
cat(draw_transitions(res_v, decision = "Intervention", output = "dot"))
```

```{r transitions-render, eval = requireNamespace("DiagrammeR", quietly = TRUE)}
draw_transitions(res_v, decision = "Intervention")
```

## Cost-effectiveness plane and CEAC

Run a probabilistic analysis (a `data.frame` of parameter draws) and visualize
the joint distribution of costs and effects with `plot_ce()`, and the
willingness-to-pay sweep with `plot_ceac()`.

```{r psa}
psa <- data.frame(rr = rbeta(100, 5, 5))
res <- run_twig(mytwig, params = psa, n_cycles = 30, progress_bar = FALSE)

plot_ce(res$sim_ev, col_names = c("cost", "qaly"))
plot_ceac(res$sim_ev, wtp_range = seq(0, 100000, by = 5000),
          col_names = c("cost", "qaly"))
```

## Decision trees

`draw_decision_tree()` visualizes the event (chance-node) structure of a
decision tree.

```{r tree-dot}
dtree <- twig() +
  decisions(names = c(Treat, DoNotTreat)) +
  event(name = response, options = c(yes, none),
        probs = c(pResponse, leftover), transitions = c(better, worse)) +
  payoffs(names = c(cost))

cat(draw_decision_tree(dtree, output = "dot"))
```

```{r tree-render, eval = requireNamespace("DiagrammeR", quietly = TRUE)}
draw_decision_tree(dtree)
```
